---
title: "Exercise solutions: Section 3.7"
author: "Rob J Hyndman & George Athanasopoulos"
output: html_document
---
```{r, echo = F, message = F}
require(fpp2)
#setwd("C:/George/fppsolutions")
```

1. For the following series, find an appropriate Box-Cox transformation in order to stabilize the variance.

    * `usnetelec`
    * `usgdp`
    * `mcopper`
    * `enplanements`


1. Use the help menu to explore what the series `gold`, `woolyrnq` and `gas` represent. These are available in the `forecast` package.

    ```{r, echo=FALSE, cache=TRUE}
    help(gold)
    help(woolyrnq)
    help(gas)
    ``` 

    a.  Use `autoplot` to plot each of these in separate plots.
    
    ```{r, echo=TRUE, cache=TRUE}
    autoplot(gold)
    autoplot(woolyrnq)
    autoplot(gas)
    ``` 
    
    b.  What is the frequency of each commodity series? Hint: apply     the `frequency()` function.
    
    ```{r, echo=TRUE, cache=TRUE}
    frequency(gold)
    frequency(woolyrnq)
    frequency(gas)
    ``` 
      
    c.  Use `which.max()` to spot the outlier in the `gold` series. Which observation was it? 

    ```{r, echo=TRUE, cache=TRUE}
    which.max(gold)
    ``` 

2. Download the file `tute1.csv` from [OTexts.org/fpp2/extrafiles/tute1.csv](extrafiles/tute1.csv) open it in Excel (or some other spreadsheet application) and review its contents. You should find four columns of information. Columns B through D each contain a quarterly series, labelled Sales, AdBudget and GDP. Sales contains the quarterly sales for a small company over the period 1981-2005. AdBudget is the advertising budget and GDP is the gross domestic product. All series have been adjusted for inflation.

    a.  You can read the data into R with the following script:

    ```{r, echo=TRUE, cache=TRUE}
    tute1 <- read.csv("tute1.csv", header=TRUE)
    #View(tute1)
    ```

    b.  Convert the data to time series

    ```{r, echo=TRUE, cache=TRUE}
    mytimeseries <- ts(tute1[,-1], start=1981, frequency=4)
    ```

    (The `[,-1]` removes the first column which contains the quarters as we don't need them now.)

    c.  Construct time series plots of each of the three series
    
    ```{r, echo=TRUE, cache=TRUE}
    autoplot(mytimeseries, facets=TRUE)
    ```

        Check what happens when you don't include `facets=TRUE`.

    ```{r, echo=TRUE, cache=TRUE, dependson="exmyts"}
    autoplot(mytimeseries)
    ``` 

